---
title: "Health Care Security Attacks"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: simplex
---

```{r setup, include=FALSE}
# Libraries

library(flexdashboard)
library(shinyWidgets)
library(shiny)
library(tidyverse)
library(DT)
library(reader)
library(rsconnect)
library(dplyr)
library(leaflet)
library(tigris)
library(sf)
library(censusapi)
library(tidycensus)
library(lubridate)
library(reactable)
library(plotly)
library(rvest)
library(highcharter)
library(dplyr)
library(viridisLite)
library(forecast)
library(flexdashboard)
library(car)
library(ggpubr)


```


```{r global, include = FALSE}
#total of 5,847 cases
all_health_breaches <- read_csv("https://raw.githubusercontent.com/Mal-Shan/HealthSecurity/main/all_health_breaches23.csv")

#reduced to 5,823 cases for missing categorical variables 
all_health_breaches <- all_health_breaches %>%
  rename("entity" = "Name of Covered Entity",
         "state" = "State",
         "type" = "Covered Entity Type",
         "affected" = "Individuals Affected",
         "date" = "Breach Submission Date",
         "incident" = "Type of Breach",
         "location" = "Location of Breached Information",
         "business_associate" = "Business Associate Present")

#collecting latest 100 cases from HHS 
url <- "https://ocrportal.hhs.gov/ocr/breach/breach_report.jsf?page=2"
file <- read_html(url)
tables <- html_nodes(file, "table")
new_data <- html_table(tables[[2]], fill = TRUE)

#selecting only the columns from the newly loaded data that match the archived data
new_data <- new_data %>%
  rename("entity" = "Name of Covered Entity",
         "state" = "State",
         "type" = "Covered Entity Type",
         "affected" = "Individuals Affected",
         "date" = "Breach Submission Date",
         "incident" = "Type of Breach",
         "location" = "Location of Breached Information",
         "business_associate" = "Business Associate Present") %>%
    select(colnames(all_health_breaches))

#Check if you need to update data with latest information
different <- generics::setdiff(new_data, all_health_breaches)
all_health_breaches <- rbind(different, all_health_breaches)

all_health_breaches <- all_health_breaches %>%   
  filter(affected < 15000000) %>%
  mutate(location = stringr::str_remove(location, ',.*'),
         incident = stringr::str_remove(incident, ',.*'),
         type = stringr::str_remove(type, ',.*')) %>%
  drop_na(c(entity, state, type, date, incident, location, business_associate))

#replacing NA values for affected values with the mean - imputation
missing_affected <- all_health_breaches[is.na(all_health_breaches$affected), ]

for(i in nrow(missing_affected)){
  
  #figure out which state the missing value came from 
  state_name <- toString(missing_affected[i,"state"])
  
  #take the average number of affected individuals that do not include this case 
  replaced_value <- all_health_breaches %>%
    drop_na(affected) %>%
    filter(state == state_name) %>%
    summarize(replaced_value = mean(affected))
  
  #replace the NA value with this mean value 
  replaced_value <- as.numeric(replaced_value)
  replaced_value <- round(replaced_value, digits = 0)
  
  replaced_entity <- toString(missing_affected[i,"entity"])
  replaced_date <- toString(missing_affected[i,"date"])
  replaced_id <- as.numeric(which(all_health_breaches$entity == replaced_entity & all_health_breaches$date == replaced_date, 
        arr.ind=TRUE))
  
  all_health_breaches[replaced_id,"affected"] <- replaced_value
  

  
}

all_health_breaches_transformed <- all_health_breaches
all_health_breaches_transformed$affected_transformed <- log10(all_health_breaches$affected)

#Creating states chloropleth dataframe
states <- states()

states <- states %>%
  sf::st_transform('+proj=longlat +datum=WGS84')

api_key <- "7fafa1056e554ba1b132c71721d21d15a938f866"

#not removing missing values
total_state_incidents <-  all_health_breaches %>%
  group_by(state) %>%
  summarize(total_affected = sum(affected), #latest addition to the total state incidents: 1/1/24
            number_attacks = n())

options(tigris_use_cache = TRUE)

state_pop <-  getCensus(name="acs/acs5", 
                        vintage=2020, #was set to 2015 before 
                        key=api_key, 
                        vars=c("NAME", "B01003_001E"), 
                        region="state:*")

colnames(state_pop) <- c("state_id", "NAME", "population")
state_pop$state_id <- as.numeric(state_pop$state_id)

state_off <- data.frame(state.abb, state.name)
colnames(state_off) <- c("state", "NAME")

state_pop <- left_join(state_pop, state_off)
state_pop$state <- ifelse(state_pop$NAME=="District of Columbia", "DC", as.character(state_pop$state))
state_pop$state <- ifelse(state_pop$NAME=="Puerto Rico", "PR", as.character(state_pop$state))

cyber_state_pop <- left_join(total_state_incidents, state_pop)

cyber_state_pop$per_capita <- round(cyber_state_pop$total_affected/cyber_state_pop$population*100000,2) #per 100,000

cyber_state_pop <- filter(cyber_state_pop, !is.na(per_capita))

states_merged_cyber_pc <- geo_join(states, cyber_state_pop, "STUSPS", "state")

pal_cyber <- colorNumeric("Reds", domain=states_merged_cyber_pc$per_capita)

states_merged_cyber_pc <- subset(states_merged_cyber_pc, !is.na(per_capita))

popup_cyber <- paste("<b>", states_merged_cyber_pc$NAME.x, 
                   "</b><br />Total Individuals Affected: ", states_merged_cyber_pc$total,
                   "<br />Per capita: ", 
                   as.character(states_merged_cyber_pc$per_capita))

dropped_total_state_incidents <-  all_health_breaches %>%
  group_by(state) %>%
  summarize(total_affected = sum(affected))

options(tigris_use_cache = TRUE)

state_pop_2 <-  getCensus(name="acs/acs5", 
                        vintage=2020, #was set to 2015 before 
                        key=api_key, 
                        vars=c("NAME", "B01003_001E"), 
                        region="state:*")

colnames(state_pop_2) <- c("state_id", "NAME", "population")
state_pop_2$state_id <- as.numeric(state_pop_2$state_id)

state_off_2 <- data.frame(state.abb, state.name)
colnames(state_off_2) <- c("state", "NAME")

state_pop_2 <- left_join(state_pop, state_off)
state_pop_2$state <- ifelse(state_pop_2$NAME=="District of Columbia", "DC", as.character(state_pop$state))
state_pop_2$state <- ifelse(state_pop_2$NAME=="Puerto Rico", "PR", as.character(state_pop$state))

cyber_state_pop_2 <- left_join(dropped_total_state_incidents, state_pop_2)

cyber_state_pop_2$per_capita <- round(cyber_state_pop_2$total_affected/cyber_state_pop_2$population*100000,2) #per 100,000

cyber_state_pop_2 <- filter(cyber_state_pop_2, !is.na(per_capita))

states_merged_cyber_pc_2 <- geo_join(states, cyber_state_pop_2, "STUSPS", "state")

pal_cyber_2 <- colorNumeric("Reds", domain=states_merged_cyber_pc_2$per_capita)

states_merged_cyber_pc_2 <- subset(states_merged_cyber_pc_2, !is.na(per_capita))

popup_cyber_2 <- paste("<b>", states_merged_cyber_pc_2$NAME.x, 
                   "</b><br />Total Individuals Affected: ", states_merged_cyber_pc_2$total,
                   "<br />Per capita: ", 
                   as.character(states_merged_cyber_pc_2$per_capita))



#counts for entity 
counts_entity <- all_health_breaches %>%
  count(entity) %>%
  rename(Attacks = n)

#updated combined attacks dataframe 

summarize_and_count <- function(df, var_interest, name){
  
  summarize_df <- df %>%
    group_by({{var_interest}}) %>%
    summarize(total_affected = sum(affected))
  
  count_df <- df %>%
    group_by({{var_interest}}) %>% 
    count({{var_interest}}) %>%
    rename(num_category = n)
    
 summarize_count <- 
   full_join(summarize_df, count_df) %>%
   rename(category = {{var_interest}}) %>%
   add_column(choice = name)

 return(summarize_count)
 
}

combined_attacks <- rbind(summarize_and_count(all_health_breaches, 
                                              type, "type"),
                          summarize_and_count(all_health_breaches, 
                                              location, "location"),
                          summarize_and_count(all_health_breaches, 
                                              incident, "incident"),
                          summarize_and_count(all_health_breaches, 
                                              business_associate, "business_associate"))


#date of attacks (year)

total_year_attacks <- all_health_breaches %>%
  mutate(date = mdy(date)) %>%
  mutate(date = ymd(paste(date, sep = " ")))  %>%
  mutate(year = year(date)) %>%
  add_count(entity) %>%
  select(state, year, n, affected) %>%
  group_by(state, year) %>%
  summarize(total_attacks = sum(n),
            total_affected = sum(affected)) #updated 1/1/24 to include the total_affected

  
#frequency of attacks
freq_attacks <- all_health_breaches %>%
  mutate(date = mdy(date)) %>%
  mutate(date = ymd(paste(date, sep = " ")))  %>%
  select(-business_associate)

#important information from merged state dataframe
sum_merged_cyber_pc <- states_merged_cyber_pc %>%
  as.data.frame() %>%
  select(NAME.x, total_affected, per_capita)


#games howell post-hoc test function definition 
games.howell <- function(grp, obs) {
  
  #Create combinations
  combs <- combn(unique(grp), 2)
  
  # Statistics that will be used throughout the calculations:
  # n = sample size of each group
  # groups = number of groups in data
  # Mean = means of each group sample
  # std = variance of each group sample
  n <- tapply(obs, grp, length)
  groups <- length(tapply(obs, grp, length))
  Mean <- tapply(obs, grp, mean)
  std <- tapply(obs, grp, var)
  
  statistics <- lapply(1:ncol(combs), function(x) {
    
    mean.diff <- Mean[combs[2,x]] - Mean[combs[1,x]]
    
    #t-values
    t <- abs(Mean[combs[1,x]] - Mean[combs[2,x]]) / sqrt((std[combs[1,x]] / n[combs[1,x]]) + (std[combs[2,x]] / n[combs[2,x]]))
    
    # Degrees of Freedom
    df <- (std[combs[1,x]] / n[combs[1,x]] + std[combs[2,x]] / n[combs[2,x]])^2 / # Numerator Degrees of Freedom
            ((std[combs[1,x]] / n[combs[1,x]])^2 / (n[combs[1,x]] - 1) + # Part 1 of Denominator Degrees of Freedom 
              (std[combs[2,x]] / n[combs[2,x]])^2 / (n[combs[2,x]] - 1)) # Part 2 of Denominator Degrees of Freedom
    
    #p-values
    p <- ptukey(t * sqrt(2), groups, df, lower.tail = FALSE)
    
    # Sigma standard error
    se <- sqrt(0.5 * (std[combs[1,x]] / n[combs[1,x]] + std[combs[2,x]] / n[combs[2,x]]))
          
    # Upper Confidence Limit
    upper.conf <- lapply(1:ncol(combs), function(x) {
      mean.diff + qtukey(p = 0.95, nmeans = groups, df = df) * se
    })[[1]]
    
    # Lower Confidence Limit
    lower.conf <- lapply(1:ncol(combs), function(x) {
      mean.diff - qtukey(p = 0.95, nmeans = groups, df = df) * se
    })[[1]]
    
    # Group Combinations
    grp.comb <- paste(combs[1,x], ':', combs[2,x])
    
    # Collect all statistics into list
    stats <- list(grp.comb, mean.diff, se, t, df, p, upper.conf, lower.conf)
  })
  
  # Unlist statistics collected earlier
  stats.unlisted <- lapply(statistics, function(x) {
    unlist(x)
  })
  
  # Create dataframe from flattened list
  results <- data.frame(matrix(unlist(stats.unlisted), nrow = length(stats.unlisted), byrow=TRUE))
  
  # Select columns set as factors that should be numeric and change with as.numeric
  results[c(2, 3:ncol(results))] <- round(as.numeric(as.matrix(results[c(2, 3:ncol(results))])), digits = 3)
  
  # Rename data frame columns
  colnames(results) <- c('groups', 'Mean Difference', 'Standard Error', 't', 'df', 'p', 'upper limit', 'lower limit')

  return(results)
}

```

Explore DHHS Dataset {data-icon="fa-table"}
===================================== 

Row {.sidebar data-width=500}
-------------------------------------

#### Real-Time Monitoring of Health Security Attacks with R-based Data Visualization Dashboard

Created By: Mridula (Mally) Shan

According to the [US Department of Health and Human Services (DHHS)](https://www.hhs.gov/sites/default/files/2020-hph-cybersecurty-retrospective-tlpwhite.pdf), in almost every month of 2020, more than 1 million people were affected by data breaches at healthcare organizations. This was a 9851% increase from 2019, leading to security exploitations in 560 total healthcare organizations. To foster discussion around these issues, I developed an interactive data dashboard using R that would allow patients and healthcare providers to fully understand the magnitude of such incidents. The data for this dashboard is continuously updated from public domain data provided by [the DHHS](https://ocrportal.hhs.gov/ocr/breach/breach_report.jsf). This project has been presented at the Association for Computing Machiner's 14th International Conference for Computational Biology, Bioinformatics, and Health informatics. Read the [ACM-BCB Conference Paper here ](https://dl.acm.org/doi/10.1145/3584371.3613035).

Use the controls below to explore the DHHS dataset:

```{r}
# From shinyWidgets

pickerInput("select_State", label = h5("Filter by Geographical Location"), 
    choices = list("Alabama" = "AL",
                   "Alaska" = "AK",
                   "Arizona" = "AZ",
                   "Arkansas" = "AR",
                   "California" = "CA",
                   "Colorado" = "CO",
                   "Connecticut" = "CT",
                   "Delaware" = "DE",
                   "District of Columbia" = "DC",
                   "Florida" = "FL",
                   "Georgia" = "GA",
                   "Hawaii" = "HI",
                   "Idaho" = "ID",
                   "Illinois" = "IL",
                   "Indiana" = "IN",
                   "Iowa" = "IA",
                   "Kansas" = "KS",
                   "Kentucky" = "KY",
                   "Louisiana" = "LA",
                   "Maine" = "ME",
                   "Maryland" = "MD",
                   "Massachusetts" = "MA",
                   "Michigan" = "MI",
                   "Minnesota" = "MN",
                   "Mississippi" = "MS",
                   "Missouri" = "MO",
                   "Montana" = "MT",
                   "Nebraska" = "NE",
                   "Nevada" = "NV",
                   "New Hampshire" = "NH",
                   "New Jersey" = "NJ",
                   "New Mexico" = "NM",
                   "New York" = "NY",
                   "North Carolina" = "NC",
                   "North Dakota" = "ND",
                   "Ohio" = "OH",
                   "Oklahoma" = "OK",
                   "Oregon" = "OR",
                   "Pennsylvania" = "PA",
                   "Puerto Rico" = "PR",
                   "Rhode Island" = "RI",
                   "South Carolina" = "SC",
                   "South Dakota" = "SD",
                   "Tennessee" = "TN",
                   "Texas" = "TX",
                   "Utah" = "UT",
                   "Vermont" = "VT",
                   "Virginia" = "VA",
                   "Washington" = "WA",
                   "West Virginia" = "WV",
                   "Wisconsin" = "WI",
                   "Wyoming" = "WY"), 
    selected = "TX", 
    multiple = TRUE, options = list(`actions-box` = TRUE, `deselect-all-text` = "Deselect All", `select-all-text` = "Select All", `none-selected-text` = "Please Make a Selection"))


pickerInput("select_Location", label = h5("Filter by Digital Location"), 
    choices = list("Desktop Computer" = "Desktop Computer", 
                   "Electronic Medical Record" = "Electronic Medical Record", 
                   "Email" = "Email",
                   "Laptop" = "Laptop",
                   "Network Server" = "Network Server",
                   "Other Electronic Device" = "Other Portable Electronic Device",
                   "Other" = "Other"), multiple = TRUE, selected = "Desktop Computer", options = list(`actions-box` = TRUE, `deselect-all-text` = "Deselect All", `select-all-text` = "Select All", `none-selected-text` = "Please Make a Selection"))


pickerInput("select_Type", label = h5("Filter by Target Entity"), 
    choices = list("Healthcare Provider" = "Healthcare Provider", 
                   "Health Plan" = "Health Plan", 
                   "Business Associate" = "Business Associate",
                   "Healthcare Clearing House" = "Healthcare Clearing House"), multiple = TRUE, selected = "Healthcare Provider", options = list(`actions-box` = TRUE, `deselect-all-text` = "Deselect All", `select-all-text` = "Select All", `none-selected-text` = "Please Make a Selection"))

pickerInput("select_Incident", label = h5("Filter by Incident Type"), 
    choices = list("Hacking/IT Incident" = "Hacking/IT Incident", 
                   "Improper Disposal" = "Improper Disposal", 
                   "Loss" = "Loss",
                   "Other" = "Other",
                   "Theft" = "Theft",
                   "Unauthorized Access" = "Unauthorized Access/Disclosure", "Unknown" = "Unknown"), multiple = TRUE, selected = "Unauthorized Access/Disclosure", options = list(`actions-box` = TRUE, `deselect-all-text` = "Deselect All", `select-all-text` = "Select All", `none-selected-text` = "Please Make a Selection"))



fluidRow(
    align = "left",
    br(),
    column(
      12,
      splitLayout(cellWidths = c("60%", "60%"),
                  actionButton(
                    inputId = "apply",
                    label = "Apply Changes",
                    style = "width:125px"
                  ),
                  actionButton(
                    inputId = "revert",
                    label = "Revert",
                    style = "width:80px"
                  )
      )
    )
  ) # fluidRow

p(" ")
p(" ")

strong("Important Note:")

p("The US Secretary of HHS is only required to report attacks in which 500 or more individuals were affected and so there may be underreporting present in this dataset")
      
```

Row
-------------------------------------

### Number of Affected Individuals this year 

```{r}

yearly_affected <- all_health_breaches %>%
  mutate(date = mdy(date)) %>%
  mutate(date = ymd(paste(date, sep = " ")))  %>%
  mutate(year = year(date)) %>%
  slice_max(year) %>%
  summarize(affected_this_year = sum(affected, na.rm = FALSE))

valueBox(yearly_affected, icon = "fa-robot", color = "darkred")
```

### Number of Attacks this month 

```{r}

monthly_attacks <- all_health_breaches %>%
  mutate(date = mdy(date)) %>%
  mutate(date = ymd(paste(date, sep = " ")))  %>%
  mutate(month = month(date),
         year = year(date)) %>%
  slice_max(year) %>%
  slice_max(month)
  
monthly_attacks <- as.numeric(nrow(monthly_attacks))

valueBox(monthly_attacks, icon = "fa-laptop-medical", color = "darkred")

```

### Number of Affected Individuals last month  

```{r}

monthly_affected <- all_health_breaches %>%
  mutate(date = mdy(date)) %>%
  mutate(date = ymd(paste(date, sep = " ")))  %>%
  mutate(month = month(date),
         year = year(date)) %>%
  filter(year == 2023) %>%
  slice_max(month) %>%
  summarize(affected_this_month = sum(affected, na.rm = TRUE))

monthly_affected <- as.numeric(monthly_affected)

valueBox(monthly_affected, icon = "fa-bugs", color = "darkred")

```


Column {.tabset data-width=500}
-------------------------------------

### History of Attacks
    
```{r}

  filteredData <- reactiveVal(freq_attacks)
  
  observeEvent(input$apply, {
    filteredData(
      freq_attacks %>% 
        filter(state %in% input$select_State, 
               type %in% input$select_Type, 
               incident %in% input$select_Incident,
               location %in% input$select_Location)
    )
      
  })
  
  observeEvent(input$revert, {
    filteredData(
      freq_attacks
    )
      
  })

  renderDataTable(datatable(filteredData()))
  
```


Map of Attacks {data-icon="fa-map"}
=====================================

Row {.sidebar data-width=300}
-------------------------------------

#### Map of Attacks

This page includes information regarding the spread of cyber-health attacks across the country. There are **TWO maps**, which are shaded based on two distinct values: the total *number of affected* individuals and the *per capita* number of affected individuals. Specifically, the per capita value was calculated by dividing the total number of affected individuals in a given state by the respective population of that state, based on [2021 Census data](https://www.census.gov/data/what-is-data-census-gov/latest-releases/2021.html). 

When comparing the total number of attacks against the per capita value of attacks, we see that more populous states **(like California, Texas, Florida and New York)** tend to have a larger number of affected individuals, but when we adjust for the respective states' populations, we see that states like **Tennessee, Delaware, and Nevada** have larger portions of their populations affected by these cyberattacks. 

Below the map is a summary table which provides a breakdown of each state's population, total number of affected residents, per capita value of affected residents, and the total number of attacks that occurred in that state. Regarding the latter value, we see again that states with **larger populations** are more likely to have a **larger number of attacks**. 

To explore these trends for yourself, you can use your mouse to hover a state and reveal a pop-up with information about the number of affected individuals from each state since 2009, along with 2021 Census data regarding each state's population. 

Row {.tabset data-width=700 data-height=600}
-------------------------------------

### Total Affected from Health Cyberattacks per State

```{r}

states_merged_cyber_pc$hover <- 
  with(states_merged_cyber_pc, paste("Total Affected", "<br>", "<br>", population, 
                                     "<br>", "State Population"))

# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)

# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white'))


fig <- plot_geo(states_merged_cyber_pc, locationmode = 'USA-states')

fig <- fig %>% add_trace(
    z = ~total_affected, locations = ~STUSPS, text = ~hover,
    color = ~total_affected, colors = 'Reds'
  )

fig <- fig %>% layout(geo = g,
                      dragmode=FALSE)

renderPlotly({
  fig
})

```  

### Per Capita Affected from Health Cyberattacks per State

```{r}

states_merged_cyber_pc$hover2 <- 
  with(states_merged_cyber_pc, paste("Per Capita Affected", "<br>", "<br>", population, 
                                     "<br>", "State Population"))

fig2 <- plot_geo(states_merged_cyber_pc, locationmode = 'USA-states')

fig2 <- fig2 %>% add_trace(text = ~hover2, z = ~per_capita, locations = ~STUSPS, text = ~hover,
    color = ~per_capita, colors = 'Reds'
  )

fig2 <- fig2 %>% layout(geo = g,
                        dragmode=FALSE)

fig2 <- fig2 %>% colorbar(title = "Per Capita Affected")

renderPlotly({
  fig2
})


```


Row {.tabset data-width=300 data-height=400}
-------------------------------------

### Summary Table

```{r}

states_info <- reactive({ 
  cyber_state_pop %>%
    select(NAME, population, number_attacks, total_affected, per_capita)
  })

renderDT({
    datatable(states_info(), 
              rownames = FALSE,
              options = list(paging = FALSE,
                             searching = FALSE,
                             orderClasses = FALSE,
                             info = FALSE))
  })
```

Comparing Multiple States {data-icon="fa-signal"}
=====================================  

Inputs {.sidebar data-width=300}
-----------------------------------------------------------------------

#### Attacks Over Time

This page allows you to further explore the data from the previous maps and compare trends of attacks overtime and across different states. There are **TWO line graphs**, with the first graph plotting the number of attacks and the second plotting the number of affected individuals, across each selected state from **2009 to the present day**. 

Use the drop down feature below to select your state(s) of interest. You can then use your mouse to hover over the graph and compare the peaks and dips of the line graphs across different states. 

```{r}

pickerInput(inputId = "filter_State2", label = h5("Select state(s) to compare:"), 
    choices = list("Alabama" = "AL",
                   "Alaska" = "AK",
                   "Arizona" = "AZ",
                   "Arkansas" = "AR",
                   "California" = "CA",
                   "Colorado" = "CO",
                   "Connecticut" = "CT",
                   "Delaware" = "DE",
                   "District of Columbia" = "DC",
                   "Florida" = "FL",
                   "Georgia" = "GA",
                   "Hawaii" = "HI",
                   "Idaho" = "ID",
                   "Illinois" = "IL",
                   "Indiana" = "IN",
                   "Iowa" = "IA",
                   "Kansas" = "KS",
                   "Kentucky" = "KY",
                   "Louisiana" = "LA",
                   "Maine" = "ME",
                   "Maryland" = "MD",
                   "Massachusetts" = "MA",
                   "Michigan" = "MI",
                   "Minnesota" = "MN",
                   "Mississippi" = "MS",
                   "Missouri" = "MO",
                   "Montana" = "MT",
                   "Nebraska" = "NE",
                   "Nevada" = "NV",
                   "New Hampshire" = "NH",
                   "New Jersey" = "NJ",
                   "New Mexico" = "NM",
                   "New York" = "NY",
                   "North Carolina" = "NC",
                   "North Dakota" = "ND",
                   "Ohio" = "OH",
                   "Oklahoma" = "OK",
                   "Oregon" = "OR",
                   "Pennsylvania" = "PA",
                   "Puerto Rico" = "PR",
                   "Rhode Island" = "RI",
                   "South Carolina" = "SC",
                   "South Dakota" = "SD",
                   "Tennessee" = "TN",
                   "Texas" = "TX",
                   "Utah" = "UT",
                   "Vermont" = "VT",
                   "Virginia" = "VA",
                   "Washington" = "WA",
                   "West Virginia" = "WV",
                   "Wisconsin" = "WI",
                   "Wyoming" = "WY"),
    selected = "AL", 
    multiple = TRUE, 
    options = list(`actions-box` = TRUE, `deselect-all-text` = "Deselect All", `select-all-text` = "Select All", `none-selected-text` = "Please Make a Selection"))

p(" ")

strong("Important Note:")

p("Since the data for these graphs are being continuously updated and the year is still ongoing, the specific number of attacks/affected individuals for the year 2024 are not definitive and will change.")

```


Row {.tabset data-width=700 data-height=600}
-------------------------------------

### Number of Attacks overtime

```{r}
  
multiple_select <- reactive({
    total_year_attacks  %>%
      filter(state %in% input$filter_State2)
  })
  
renderPlotly({
  
  ggplot(data = multiple_select(), 
         mapping = aes(x = year, y = total_attacks, color = state)) +
    geom_line(lwd = 0.5,    # Line width
             linetype = 1,
             size = 3) +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_blank(), 
          plot.background = element_rect(fill = "#f5f5f5"),
          axis.line = element_line(colour = "black")) + 
    scale_x_continuous(breaks = round(seq(min(multiple_select()$year), 
                                          max(multiple_select()$year), 
                                          by = 1), 1)) + 
    labs(y = "# Attacks", x = "Year") +
    theme(plot.title = element_text(size = 15, face = "bold"),
          axis.text = element_text(size = 15),
          axis.text.x = element_text(face = "bold",
                                     angle = 45,
                                     vjust = 1,
                                     hjust = 1),
          axis.text.y = element_text(face = "bold"))

  })

```

### Number of Affected Individuals

```{r}
  
multiple_select <- reactive({
    total_year_attacks  %>%
      filter(state %in% input$filter_State2)
  })
  
renderPlotly({
  
  ggplot(data = multiple_select(), 
         mapping = aes(x = year, y = total_affected, color = state)) +
    geom_line(lwd = 0.5,    # Line width
             linetype = 1,
             size = 3) +
    theme(panel.grid.major = element_blank(), 
          panel.grid.minor = element_blank(),
          panel.background = element_blank(), 
          plot.background = element_rect(fill = "#f5f5f5"),
          axis.line = element_line(colour = "black")) + 
    scale_x_continuous(breaks = round(seq(min(multiple_select()$year), 
                                          max(multiple_select()$year), 
                                          by = 1), 1)) + 
    labs(y = "# Affected", x = "Year") +
    theme(plot.title = element_text(size = 15, face = "bold"),
          axis.text = element_text(size = 15),
          axis.text.x = element_text(face = "bold",
                                     angle = 45,
                                     vjust = 1,
                                     hjust = 1),
          axis.text.y = element_text(face = "bold")) + 
    scale_y_continuous(trans='log10')

  })

```

Row {.tabset data-width=300 data-height=400}
-------------------------------------

### Summary Table
    
```{r}

multiple_select_desc <- reactive({ 
    multiple_select() %>%
    arrange(desc(year)) %>%
    select(state, year, total_affected, total_attacks)

  })

renderDT({
    datatable(multiple_select_desc(), 
              rownames = FALSE,
              options = list(paging = FALSE,
                             searching = FALSE,
                             orderClasses = FALSE,
                             info = FALSE))
  })

```


Interactive Modeling (ANOVA) {data-icon="fa-table"}
===================================== 

Row {.sidebar data-width=500}
-------------------------------------

#### Interactive Modeling (ANOVA)

This page explores the different types of cyberattacks and seeks to determine which types of attacks lead to the most amount of affected individuals. Specifically, the DHHS examines **four different categories** of attacks:

* **incidents** (Hacking/IT Incident, Theft, Unauthorized Access/Disclosure, Loss, Improper Disposal)
* **type** of target entities (Healthcare Provider, Business Associate, Health Plan, Healthcare Clearing House)
* digital **locations** of attacks (Network Server, Email, Paper/Films, Other Portable Electronic Device, Desktop Computer, Electronic Medical Record, Laptop)
* presence or absence of **business associates**

Below, an initial [ANOVA test](https://www.qualtrics.com/experience-management/research/anova/) is run, demonstrating that these variables have an impact on the number of affected individuals for a given cyberattack, with **p<0.01** for every category, as emphasized by the number of asteriks following the Pr(>F) values. 

```{r}

print(summary(aov(affected_transformed ~ incident + type + location + business_associate, 
            data = all_health_breaches_transformed)))

```

We can then run a **one-way ANOVA test**, this time focusing on a specific variable, to see which of the sublevels of the variable are significantly different in the median number of affected individuals, as determined by a **post-hoc Tukey HSD test**. 

```{r}

variables2 <- c("incident","type", "location",  "business_associate")

radioButtons("variable_choice2", 
            label = h5("Choose a variable to explore:"),
            choices = variables2, 
            variables2[1])

p()

model2 <- reactive({
  
  if(input$variable_choice2 == "incident"){
    aov(affected_transformed ~ incident, 
                     data = all_health_breaches_transformed)
  }
  else if(input$variable_choice2 == "type"){
   aov(affected_transformed ~ type, 
                    data = all_health_breaches_transformed)
  }
  else if(input$variable_choice2 == "location"){
    aov(affected_transformed ~ location, 
                     data = all_health_breaches_transformed)
  }
  else if(input$variable_choice2 == "business_associate"){
    aov(affected_transformed ~ business_associate, 
                     data = all_health_breaches_transformed)
  }
  
})

renderPrint({ 
   
Anova(model2(), type="II")

})


```

For example, if you select the **type** variable, we see from the initial exploratory visualizations that there are **differences in the median** number of affected individuals depending on whether the attack is directed against a Healthcare Provider, Business Associate, Health Plan or Healthcare Clearing House. We see that the confidence interval for a **Tukey multiple comparisons of means** test of **Health Plan and Business Associate** ranges from -0.326 to -0.118, while **Healthcare Provider and Business Associate** ranges from -0.277 to -0.121. Since neither range includes zero, this indicates that the difference of means of affected individuals is statistically significant. The confidence intervals for the remaining pairs of means all include zero (as visually represented in the Tukey HSD graph), which indicates that the differences are not statistically significant. 


Row {.tabset data-width=500}
-------------------------------------

### Data Visualization #1

```{r}

renderPlotly({
  
  if(input$variable_choice2 == "incident"){

  ggboxplot(all_health_breaches_transformed, 
            x = "incident", 
            y = "affected_transformed", 
            color = "incident") +
      theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
  }
  
  else if(input$variable_choice2 == "type"){
    
    ggboxplot(all_health_breaches_transformed, 
            x = "type", 
            y = "affected_transformed", 
            color = "type") +
      theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
    }
  
  else if(input$variable_choice2 == "location"){

  ggboxplot(all_health_breaches_transformed, 
            x = "location", 
            y = "affected_transformed", 
            color = "location") +
      theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
    }
  
  else if(input$variable_choice2 == "business_associate"){

  ggboxplot(all_health_breaches_transformed, 
            x = "business_associate", 
            y = "affected_transformed", 
            color = "business_associate") +
      theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
  }
  
})


```

### Data Visualization #2

```{r}

renderPlotly({
  
  if(input$variable_choice2 == "incident"){

ggline(all_health_breaches_transformed, 
       x = "incident", y = "affected_transformed",  
       add = c("mean_se", "jitter")) +
  scale_y_continuous(trans='log10') + theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
    
  }
  
  else if(input$variable_choice2 == "type"){
    
ggline(all_health_breaches_transformed, 
       x = "type", y = "affected_transformed",  
       add = c("mean_se", "jitter")) +
  scale_y_continuous(trans='log10') + theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
    
    }
  
  else if(input$variable_choice2 == "location"){

ggline(all_health_breaches_transformed, 
       x = "location", y = "affected_transformed",  
       add = c("mean_se", "jitter")) +
  scale_y_continuous(trans='log10') + theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
    
    }
  
  else if(input$variable_choice2 == "business_associate"){

ggline(all_health_breaches_transformed,
       x = "business_associate", y = "affected_transformed",
       add = c("mean_se", "jitter")) +
  scale_y_continuous(trans='log10') + theme(text = element_text(size = 8), 
				axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
    
  }
  
})


```

### Tukey HSD graph

```{r}
renderPlot({
  
par(mar = c(4.1, 13.5, 4.1, 2.1))
# create confidence interval for each comparison
plot(TukeyHSD(model2()),
  las = 2 # rotate x-axis ticks
)

})

```

### Tukey HSD test 

```{r}

renderPrint({ 
  
TukeyHSD(model2())
  
})

```

### Summary Table

```{r}

dat_type_agg <- reactive({
  
  combined_attacks %>%
    filter(choice == input$variable_choice2) %>%
    select(category, num_category, total_affected)
  
  })
  
renderDT({
    datatable(dat_type_agg(),
              options = list(paging = TRUE,
                             searching = FALSE,
                             orderClasses = TRUE))
  })

```



Interactive Modeling (Welch ANOVA) {data-icon="fa-table"}
===================================== 

Row {.sidebar data-width=500}
-------------------------------------

#### Interactive Modeling (Welch ANOVA)

This page expands upon the previous exploration of the different types of cyberattacks and seeks to determine which types of attacks lead to the most amount of affected individuals. Specifically, the DHHS examines **four different categories** of attacks:

* **incidents** (Hacking/IT Incident, Theft, Unauthorized Access/Disclosure, Loss, Improper Disposal)
* **type** of target entities (Healthcare Provider, Business Associate, Health Plan, Healthcare Clearing House)
* digital **locations** of attacks (Network Server, Email, Paper/Films, Other Portable Electronic Device, Desktop Computer, Electronic Medical Record, Laptop)
* presence or absence of **business associates**

Below, an initial [ANOVA test](https://www.qualtrics.com/experience-management/research/anova/) is run, demonstrating that these variables have an impact on the number of affected individuals for a given cyberattack, with **p<0.01** for every category, as emphasized by the number of asteriks following the Pr(>F) values. 

```{r}

print(summary(aov(affected_transformed ~ incident + type + location + business_associate, 
            data = all_health_breaches_transformed)))

```

However, in order to run an ANOVA test, the following assumptions must be met: 

* the groups must have the "same" variance (homoscedasticity)
* each group must be normally distributed
* the samples must be randomly selected in an independent manner

```{r}
variables <- c("incident","type", "location",  "business_associate")

# From shinyWidgets

radioButtons("variable_choice", 
            label = h5("Choose a variable to test if it meets these conditions:" ),
            choices = variables, 
            variables[1])

p()
```

The first condition can be tested by running Levene's test. Since p<0.01 for all of the variables, we see that an ANOVA test alone is not enough evidence to demonstrate that these variables have a significant affect on the number of affected individuals. Other assumptions, such as viewing the normality of the model through a QQPlot, are demonstrated in the "Assumptions Check" tab to the right. 

```{r}

p()

levene_results <- reactive({
  
  if(input$variable_choice == "incident"){
    leveneTest(affected_transformed ~ incident, 
               all_health_breaches_transformed)
  }
  else if(input$variable_choice == "type"){
    leveneTest(affected_transformed ~ type, 
               all_health_breaches_transformed)
  }
  else if(input$variable_choice == "location"){
    leveneTest(affected_transformed ~ location, 
               all_health_breaches_transformed)
  }
  else if(input$variable_choice == "business_associate"){
    leveneTest(affected_transformed ~ business_associate, 
               all_health_breaches_transformed)
  }
  
})

renderPrint({
  levene_results()
})


```

Instead, by running an ANOVA Welch test which does not rely on the homoscedasticity assumption, we are more certain that these variables have a significant influence on the number of affected individuals. The post-hoc test for an ANOVA Welch test is Games Howell, with those values also provided on this page. 

```{r}

p()

model1 <- reactive({
  
  if(input$variable_choice == "incident"){
    oneway.test(affected_transformed ~ incident, 
                     data = all_health_breaches_transformed)
  }
  else if(input$variable_choice == "type"){
   oneway.test(affected_transformed ~ type, 
                    data = all_health_breaches_transformed)
  }
  else if(input$variable_choice == "location"){
    oneway.test(affected_transformed ~ location, 
                     data = all_health_breaches_transformed)
  }
  else if(input$variable_choice == "business_associate"){
    oneway.test(affected_transformed ~ business_associate, 
                     data = all_health_breaches_transformed)
  }
  
})

renderPrint({ 
   
model1()

})

```

Row {.tabset data-width=500}
-------------------------------------

### Assumptions Check 

```{r}

model3 <- reactive({
  
  if(input$variable_choice == "incident"){
    aov(affected_transformed ~ incident, 
                     data = all_health_breaches_transformed)
  }
  else if(input$variable_choice == "type"){
   aov(affected_transformed ~ type, 
                    data = all_health_breaches_transformed)
  }
  else if(input$variable_choice == "location"){
    aov(affected_transformed ~ location, 
                     data = all_health_breaches_transformed)
  }
  else if(input$variable_choice == "business_associate"){
    aov(affected_transformed ~ business_associate, 
                     data = all_health_breaches_transformed)
  }
  
})

renderPlot({

  par(mfrow=c(2,2))
  plot(model3())
  par(mfrow=c(1,1))

})

```


### Games Howell test

```{r}

post_hoc1 <- reactive({ 
  
   if(input$variable_choice == "incident"){
    
    games.howell(all_health_breaches_transformed$incident,
                 all_health_breaches_transformed$affected_transformed)
    
  }
  
  else if(input$variable_choice == "type"){
    
   games.howell(all_health_breaches_transformed$type,
                 all_health_breaches_transformed$affected_transformed)
    
  }
  
  else if(input$variable_choice == "location"){
    
   games.howell(all_health_breaches_transformed$location,
                 all_health_breaches_transformed$affected_transformed)
    
  }
  
  else if(input$variable_choice == "business_associate"){
    
    games.howell(all_health_breaches_transformed$business_associate,
                 all_health_breaches_transformed$affected_transformed)
    
  }
  
  })

renderDT({
    datatable(post_hoc1(), 
              rownames = FALSE,
              options = list(paging = FALSE,
                             searching = FALSE,
                             orderClasses = FALSE,
                             info = FALSE))
  })


```

